Finiteness in Cubical Type Theory

### 2020-06-16 sections/introduction.tex
more conversion to agda

  \begin{center}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
    \end{agdalisting}
  \end{center}
  We will use different notations to refer to this type depending on the
  setting.
  The following four expressions all denote the same type:

  \begin{multicols}{4} \centering
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
    \end{agdalisting}

@@

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
    \end{agdalisting}
  \end{multicols}

  The non-dependent product is a special instance of the dependent.
  We indicate a simple pair of types \(A\) and \(B\) as \(A \times B\).
\end{definition}
\begin{definition}[Dependent Product]
  Dependent products (dependent functions) use the \(\Pi\) symbol.

@@

\begin{definition}[Disjoint Union]
  We define disjoint union as an inductive type.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}
  It is also expressible with only \(\Sigma\):
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
  for definitions.
  \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
  Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
  so the distinction isn't terribly important in usage: we will only use one
  or the other as a suggestion of how we constructed it or how it is to be
  used.
\end{definition}

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{definition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  Some have higher homotopies (paths which aren't unique).
  We actually have a hierarchy of complexity of structure of path spaces in
  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
  \cite[definition 3.1.1]{hottbook}.
  \begin{alignat}{2}
    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y)
  \end{alignat}
\end{definition}
\begin{definition}[Fibres] \label{fibres}
  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{equation}
    \text{fib}_f(y) \coloneqq \Sigma(x : A) , (f (x) \equiv y)
  \end{equation}
\end{definition}
\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
  \end{alignat}
\end{definition}
\begin{definition}[Decidable Types]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{definition}
\begin{definition}[Discrete Types]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{definition}
\begin{definition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{definition}
\begin{definition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

>>>>

  \begin{center}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
    \end{agdalisting}
  \end{center}
  We will use different notations to refer to this type depending on the
  setting.
  The following four expressions all denote the same type:
  \begin{multicols}{4}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
    \end{agdalisting}

@@

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
    \end{agdalisting}
  \end{multicols}
  The non-dependent product is a special instance of the dependent.
  We indicate a simple pair of types \(A\) and \(B\) as \(A \times B\).
\end{definition}
\begin{definition}[Dependent Product]
  Dependent products (dependent functions) use the \(\Pi\) symbol.

@@

\begin{definition}[Disjoint Union]
  We define disjoint union as an inductive type.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}
  It is also expressible with only \(\Sigma\):
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equalities, equivalences, and paths]
  \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
  Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
  so the distinction isn't terribly important in usage: we will only use one
  or the other as a suggestion of how we constructed it or how it is to be
  used.
\end{definition}

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{definition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  Some have higher homotopies (paths which aren't unique).
  We actually have a hierarchy of complexity of structure of path spaces in
  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
  \cite[definition 3.1.1]{hottbook}.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
  \end{agdalisting}
\end{definition}

\begin{definition}[Fibers] \label{fibers}
  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
  \end{alignat}
\end{definition}
\begin{definition}[Decidable Types]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{definition}
\begin{definition}[Discrete Types]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{definition}
\begin{definition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{definition}
\begin{definition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).
